---
title: "Readme"
author: "Atul Gunjal"
date: "24 September 2017"
output: html_document
--- 
# Readme documnet for the week4 assignnmnet


#Raw Data
The raw data for this project is accelerometer data collected from the Samsung Galaxy S smartphone, and was provided to us at the links below:

A full description is available at the site where the data was obtained:


**Data file :** 
https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

**Full description of data :** http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones

This data included both the raw sampled data (folder ../Inertial Signals) and features apparently based on the raw data. For the purpose of this project, I am only looking at the features, not the raw data.
There are 3 types of files:

**.	x:** 
rows of feature measurements

**.	y:** 
the activity labels corresponding to each row of X. Encoded as numbers.

**.	subject:** 
the subjects on which each row of X was measured. Encoded as numbers.


In addition, to determine which features are required, we look at the list of features:

**.	features.txt**
The encoding from activity labels ids to descriptive names.

**.	activity_labels.txt** 


#Data load

**.	Setup working directory**

Set up workign directory where all input and outut file are residing.

**.	Unzip the zip file**

It creates the folder structure as follows:

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;* Sub-directory : UCI HAR Dataset


&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*Under this sub directory two folders are created

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*test

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*train


&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*Under "test" folder following files are copies. These are in .txt format.

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*subject_test

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*X_test

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*Y_test


&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*Under "train" folder following files are copies. These are in .txt format. 

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*subject_train

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*X_train

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*Y_train


.	The data from  X, y and  subject data from "test" is combined in to a dataset. It is called as *df_X__Test_Total*

.	The data from  X, y and  subject data from "train" is combined in to a dataset. It is called as *df_X__Train_Total*

.	A combined data *df_Total* is created by appending *df_X__Train_Total* to *df_X__Test_Total*

.	From the combine data a subset is created for the measures which contain the 'mean' and 'std' characters in the column names.

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*To achieve this read the "features.txt file. 

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*From this create vector "vector_meanandstdcolumnNames". It contains column numbers of those columns which contains mean and std characters in it. 

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;*Now, create a subset, dfSubsetofTotal from df_total. 

```{r echo = FALSE}
## This piece of code does seting up fo the working directory, loading required library, readting test, training and subject data. Combine the test data.


setwd("D:\\Data Science Course\\Coursera-R-Programming_Course\\Getting and Cleaning Data\\week4\\Data")
library(dplyr)
library(tidyr)
df_X__Test_Total = NULL
df_X_Train_Total = NULL
df_Total = NULL
dfSubsetofTotal = NULL
dfSubsetofTotal2 = NULL
dfSubsetofTotal3 = NULL

##Step1.1: Read X_test, Y_test and Subject_test  data 
##and merge it into one block say test_block
dfy_test <- read.table(".\\y_test.txt",header = FALSE)
dfX_test <- read.table(".\\X_test.txt",header = FALSE)
dfSubject_test <- read.table(".\\subject_test.txt",header = FALSE)
df_X__Test_Total <- cbind(dfSubject_test,dfy_test,dfX_test)

##Step1.2: Read X_train, Y_train and Subject_train  data
##and merge it into one block say train_block
dfSubject_train <- read.table(".\\subject_train.txt",header = FALSE)
dfy_train <- read.table(".\\y_train.txt",header = FALSE)
dfX_train <- read.table(".\\X_train.txt",header = FALSE)
df_X_Train_Total <- cbind(dfSubject_train,dfy_train,dfX_train)

##Step 1.3 Combine test_block and train_block into one 
df_Total <- rbind(df_X__Test_Total, df_X_Train_Total)


```


#Transformation

1.	Now we have a data frame, *dfSubsetofTotal*. The datafram is subset of the *df_Total*. It contains the data of those columns whihc has characters 'mean' and 'std' in their column names. 

&nbsp;1.1	We have to make data frame, *dfSubsetofTotal*, column names more human readable. FOr this we need to replace atcivity id column by by activity description. To achieve this we take following steps. 

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;1.1.1.	Read feature.txt file. Create a dataframe, *dfFeature*.  Create a vecor from it, *vector_meanandstdcolumnNames*, by selecting the column numbers which have  characters 'std' and 'mean' embed with in it. 

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;1.1.2.	Referring to the data frame, *df_Total* and vecor,*vector_meanandstdcolumnNames*, crate the data frame *dfSubsetofTotal* .

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;1.1.3.	Create a vector, *colNAmestossign*, by referring to the vector, *vector_meanandstdcolumnNames*, and the data frame *dfFeature*.  This vector contains the name to assign to the columns to make data more readable. 

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;1.1.4.	From the vector, *colNAmestossign*, change some column names by making it more readable. i.e. replace first "t" and "f" with "time" and "freq".  Assign these names to the dtaframe *dfSubsetofTotal*. 

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;1.1.5.	CHange first two column names of the dataframe *dfSubsetofTotal* with "Subject" and "Activity" respectively. 

1.2	Group data Subject and Activitywise from the dataframe *dfSubsetofTotal*. And create a dataframe *dfSubsetofTotal2*. 

1.3 Take a mean of each activity amd create a new dataframe *dfSubsetofTotal3*

1.4 Write the resultant dataframe,*dfSubsetofTotal3*,   as a text into the file, *MyTidyData.txt*. 
in to the 

```{r echo = FALSE}
##Step 2
##Extract only subject, activities,  and measuremnet columns 
##with mean and standarad names in it 

##Step1 2.1 read feature.txt and select columns with mean and std name
dfFeature <- read.table(".\\features.txt",header = FALSE)
vector_meanandstdcolumnNames <- grep("mean|std", dfFeature$V2)

##Step 2.2 cerate a subset of mean/std measures from the total measure 
dfSubsetofTotal <- df_Total[,vector_meanandstdcolumnNames]  
##dfSubsetofTotal[,2] <- lookUpVector[dfSubsetofTotal[,2]]
##Step 3. Change subset column names to more readable one
dfFeature$V2[ vector_meanandstdcolumnNames[1:77]]
colNAmestossign <- as.vector (dfFeature$V2[ vector_meanandstdcolumnNames])
X <- sub ("t", "time", colNAmestossign)  #replace first char "t" of the column name with time
y <- sub ("f", "freq", X) #replace first char "f" of the column name with freq

colNAmestossign <- y # more readable column names  name 


colnames(dfSubsetofTotal)[c(3:79)] <- colNAmestossign[1:77]
colnames(dfSubsetofTotal)[c(1:2)] <- c("Subject","Activity")

##Step4 remove punctuation marks from colnames 
##and substiture first t with time, and f with freq 
#gsub("[[:punct:]]","",colnames(dfSubsetofTotal)[c(3:79)])
##sub("f", "freq-",colnames(dfSubsetofTotal)[c(3:79)])

##Step 4.1. Group subset by SUbject and Activity using group_by
dfSubsetofTotal2 <- group_by(dfSubsetofTotal,Subject,Activity) 
#replace the NULL's with 0
dfSubsetofTotal2[is.null(dfSubsetofTotal2)] <- 0.0

##Step 5 mean of each measures. 
dfSubsetofTotal3 <- summarise_each(dfSubsetofTotal2,funs(mean), 3:77)

write.table(dfSubsetofTotal3, file = "MyTidyData.txt", sep="\t", row.names=FALSE)
 


```




`